### 2023-11-07 paper/paper2/ms.tex
Pair writing paper

\documentclass[twocolumn]{aastex631}
\bibliographystyle{aasjournal}

\usepackage{amsmath}
\usepackage{blindtext}
\usepackage{bm}
\usepackage{enumitem}
% Temporary packages
\usepackage{mdframed}
\usepackage{verbatim}

\def\Teff{T_{\rm eff}}
\def\vsini{v\sin{i}}
\def\kmps{\mathrm{km}\;\mathrm{s}^{-1}}

\begin{document}
\title{Blas\'e3D}
\shorttitle{\emph{blas\'e3D}}

\author[0000-0002-2290-6810]{Sujay Shankar}
\author[0000-0002-4020-3457]{Michael Gully-Santiago}
\author[0000-0002-4404-0456]{Caroline V. Morley}
\affil{Department of Astronomy, The University of Texas at Austin, 2515 Speedway, Austin, TX 78712, USA}
\shortauthors{Shankar \& Gully-Santiago \& Morley}

\begin{abstract}
    \blindtext
\end{abstract}

\keywords{}

\section{Introduction}
\blindtext

\begin{mdframed}
    \textbf{Figure: Flowchart-with-words}
\end{mdframed}

\section{Emulating the PHOENIX grid}

\begin{mdframed}
    \textbf{The PHOENIX grid subset}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Emulation with blas\'e}

    - Per-grid point optimization procedure
    - Pretrained model caching
    - Output storage

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Open Source Availability}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Figure: Line density across the grid heatmap}
\end{mdframed}

\section{Line-by-Line Fundamental Stellar Properties}
\begin{mdframed}
    \textbf{Conceptual Illustration: Faceted Plot (Teff, Logg, Z) -> Line Profiles}
\end{mdframed}

\begin{mdframed}
    \textbf{Line Recognition}
    - Need: Identifying unique lines
    - Anticipated friction points: line centroids drift
    - Strategy: (Pre-shift centers)
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Line property bulk trends across the grid}
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Figure: Heatmap $T_\mathrm{eff}$ vs $\log{g}$ for a single line $2\times2$ panels for $\sigma$, $\gamma$, $A$, $\lambda$ }
\end{mdframed}

\begin{mdframed}
    \textbf{Anomalies in Heatmaps}
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\section{Mapping Line Parameters to Fundamental Properties}
\begin{mdframed}
    \textbf{Statement: A Bidirectonal Relation exists}
    - Goal: identify a functional form
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Figure: Flowchart-with-equations}
\end{mdframed}

\begin{mdframed}
    \textbf{Functional Form}
    - Describe functional form options and refinement procedures
    - List possibilities, Linreg, GPs \citep{2023ARA&A..61..329A}, NNs, etc.
    - We choose LSTSQ
    - Enumerate functional form
    - Adapting model complexity-- AIC
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Problem: missing lines}
    - Conceivable Solution 1: treat as NaNs and deal with sparsity
    - Conceivable Solution 2: increase model complexity
    \\--------\\
    When using blas\'e, we aggregate the detected lines from all of the grid points
    into one superset of lines. However, not all of the lines in this superset are 
    detected at every grid point. In order to deal with this sparsity, we ignore lines
    have $xx$ detections or less. This removes the most sparse lines, but some sparsity still
    remains. The remaining 'missing lines' are dealt with by treating them as Voigt profiles 
    with amplitudes of 0, and Gaussian and Lorentzian widths equal to $xx$ and $xx$ respectively.
\end{mdframed}

\section{Performance evaluation}

\begin{mdframed}
    \textbf{Typical Line reconstruction performance}
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \emph{stretch goal}\par
    \textbf{End-to-end PHOENIX grid replication and residual}
    - State the per-pixel residual
    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\section{Discussion}
\begin{mdframed}
    \textbf{Revisiting Model Assumptions}

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Limitations}

    - Computational resources
    - Line profile inaccuracy
    - Surface functional form

    \textcolor{lightgray}{\blindtext}
\end{mdframed}

\begin{mdframed}
    \textbf{Conceivable Extensions}
    \\-------------\\
    Future studies can extend this work in several ways:
    \begin{enumerate}[label=-]
        \item Extending the data from just the PHOENIX Grid to additional synthethic spectral
        model grids, such as the Sonora substellar model grids or the COOLTLUSTY substellar model grid.
    \end{enumerate}
\end{mdframed}

\pagebreak
\newpage

\begin{acknowledgments}
    \blindtext
\end{acknowledgments}

\software{}

\bibliography{ms}
\clearpage

\appendix
\section{Notation}
We adopt similar notation to the blas\'e paper, with some small modifications.

\begin{deluxetable}{cp{10cm}}[h]
    \tabletypesize{\small}
    \tablecaption{Notation used in this paper\label{table2}}
    \tablehead{
        \colhead{Symbol} & \colhead{Meaning}
    }
    \startdata
    \hline
    \multicolumn{2}{c}{Indexes}\\
    \hline
    $i$ & Pixel index of the precomputed spectrum\\
    $j$ & Spectral line index\\
    $k$ & Grid point index\\
    \hline
    \multicolumn{2}{c}{Spectra}\\
    \hline
    $\bm{\lambda}$ & Spectral axis of the model grid\\
    $\bm{G}$ & The grid of synthetic spectral models\\
    $\mathsf{B}_k$ & Blackbody spectrum of the $k^{th}$ model\\
    $\mathsf{P}_k$ & Continuum polynomial of the $k^{th}$ model\\
    $\mathsf{S}_k$ & Preprocessed flux vector of the $k^{th}$ model\\
    $\mathsf{\hat{S}}$ & Flux vector of the continuous model reconstruction\\
    $\mathsf{R}_k$ & Residual between a $\mathsf{S}_k$ and its reconstruction\\
    \hline
    \multicolumn{2}{c}{Spectral Line Properties}\\
    \hline
    $\mu_{jk}$ & Center position of the $j^{th}$ line in the $k^{th}$ model\\
    $A_{jk}$ & Amplitude of the $j^{th}$ line in the $k^{th}$ model\\
    $\sigma_{jk}$ & Gaussian shape parameter of the $j^{th}$ line in the $k^{th}$ model\\
    $\gamma_{jk}$ & Lorentzian shape parameter of the $j^{th}$ line in the $k^{th}$ model\\
    \hline
    \multicolumn{2}{c}{Scalars}\\
    \hline
    $I$ & Number of pixels in the model grid's spectral axis\\
    $J$ & Number of spectral lines detected across the grid \\
    $K$ & Number of grid points in the model grid\\
    $p$ & The prominence threshold for spectral line identification\\
    $w$ & Wing cut pixel threshold for spectral line identification\\
    $v_r$ & Radial velocity of the star\\
    $R$ & Spectrograph resolving power $\lambda/\delta\lambda$\\
    $Q$ & Quality of a spectral reconstruction\\
    $H$ & The average storage impact of a spectral reconstruction\\
    $T$ & The average computational impact of a spectral reconstruction\\
    \enddata
\end{deluxetable}

\end{document}