### 2024-03-27 paper/paper2_neo/paper.tex

\documentclass[twocolumn]{aastex631}
\usepackage{graphicx}
\usepackage{physics}
\usepackage[super]{nth}
\usepackage{enumitem}

\begin{document}
\title{Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference}
\author[0000-0002-2290-6810]{Sujay Shankar}
\author[0000-0002-4020-3457]{Michael Gully-Santiago}
\author[0000-0002-4404-0456]{Caroline V. Morley}
\affil{Department of Astronomy, The University of Texas at Austin, Austin, TX 78712, USA}

\begin{abstract}
    Text
\end{abstract}
\keywords{Text}

\section{Introduction}
Text

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the
PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
the grid, whose parameter ranges are given in Table 1.

\begin{table}[h]
    \hspace*{-1.1cm}\begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\textbf{PHOENIX Grid Subset}}\\
        \hline\hline
        Parameter & Symbol & Range\\
        \hline
        Alpha Element Abundance & $\alpha$ & [0] dex\\
        Iron Abundance & [Fe/H] & [-0.5, 0.5] dex\\
        Effective Temperature & $T_{\mathrm{eff}}$ & [2300, 12000] K\\
        Surface Gravity & $\log(g)$ & [0, 5]\\
        Wavelength & $\lambda$ & [8050, 12850] \AA\\
        \hline
    \end{tabular}
    \caption{The subset of the PHOENIX grid used in this study. These limits 
    were imposed to reduce the computational cost of the algorithms. The 
    wavelength limits in particular roughly line up with that of the 
    Habitable Zone Planet Finder (HPF) spectrograph.}
\end{table}

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was accessed directly from the PHOENIX website
using \texttt{gollum}'s download option, which uses the website's FTP server
to download spectra without using any disk space. This came at the cost of
longer load times due to internet connection being significantly more of a
bottleneck than disk I/O, but was acceptable as this was a one-time operation.

The spectra were then put through a three-step preprocessing pipeline.
\begin{enumerate}
    \item Blackbody Division: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item Percentile Normalization: The spectra were normalized by dividing
    them by their 99th percentile.
    \item Continuum Normalization: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into an interpretable format.
We wanted to represent the spectra as a list of spectral lines rather than a
list of fluxes. This was done using \texttt{blase}, which detects spectral 
lines as Voigt profiles and tunes the profiles to mimic the original 
PHOENIX spectrum with back propagation. Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating small numbers. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s line detection and optimization at the expense of an
increased computational cost.

Once optimization was complete, the list of identified lines, present in the
state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
Earlier we mentioned that \texttt{blase} tuned the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset. We decided to do
this by using the line centers of the detected lines pre-optimization. This
introduced a new variable called the jitter $\delta\mu$, which measures the shift of the
line center between pre-optimization and post-optimization. Now with each
spectral line indexed by $\mu$, we had four parameters to interpolate:
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$.

There was also a second issue, and that is that spectral lines were often
only detected in some spectra from the PHOENIX subset. This means that in 
theory, interpolation would be inaccurate in regions where a line does not 
appear, and also breaks in practice because \texttt{scipy}'s regular grid
interpolation relies on a rectilinear grid, which we would not have if
regions of grid points were missing. To solve this, we artificially
populated missing grid points with log-amplitudes of -1000, which retained
interpolator stability by not being an infinity, but also essentially 
nullified the line when evaluated.

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and $[Fe/H]$, and the output was a list of four parameters, 
$\delta\mu$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators represent multiple 
manifolds mapping fundamental stellar parameters to spectral line properties. 
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 19 GB of disk space.

\section{Bayesian Optimization}
Text

\section{Inference Testing}
\subsection{Test Data}
Text

\subsection{Performance Results}
Text

\section{Discussion}
It should be noted that this study represents a proof of concept, and that 
there are numerous design considerations that could be improved upon with 
future work. These considerations include but are not limited to the 
following:
\begin{enumerate}[label=-]
    \item \textit{Limited PHOENIX Subset}: The PHOENIX subset used in this 
    study was just that, a subset. The full PHOENIX grid not only expands 
    the [Fe/H] range to [-4.0, 1.0] dex, but also includes the alpha element 
    abundance, which we elected to fix at 0 for this study. In addition to 
    the actual fundamental stellar parameters, we also took a subset of the 
    PHOENIX wavelength range, with the full [500, 55000] \AA \ wavelength 
    range also being left to future work.
    \item \textit{Strict Wavelength Range}: Currently, the generator only 
    supports inference on spectra whose wavelength limits are either equal 
    to it, or encompass that of the generator and have been truncated to 
    match. However, when the spectrum in question has a smaller wavelength 
    range than the generator, currently there is no functionality to truncate 
    the generator. This would require externally indexing the generator's 
    individual interpolators by line center position and selectively
    evaluating those to eliminate wasteful computation.
    \item \textit{Single Model Grid}: The PHOENIX grid is not the only model
    grid of synthetic spectra available, and it does not apply to all types 
    of stars. Future work would extend the reach of this study's algorithm 
    to encompass other model grids such as the Sonora series for substellar
    models. \texttt{blase} should be able to have an option for the user to 
    input which model grid they would like to base the inference on, and 
    to get even more advanced, perhaps even have the ability to intelligently
    determine which model grid to use automatically.
    \item \textit{Memorization vs.\ Generalization}: The current design of
    the algorithm constructs manifolds using interpolation. This means that
    performance is good at points close to PHOENIX subset grid points, but
    is highly dependent on the type of interpolation used. As interpolators
    require memorization of the data, advanced interpolation becomes extremely
    expensive in terms of disk utilization. Future work would involve 
    constructing manifolds using regression, which would allow for much
    better generalization and lower disk utilization at the expense of
    some accuracy.
    \item \textit{Extrinsic Absence}: The current design of our algorithm
    does not account for extrinsic parameters that modify the appearance of 
    spectra such as rotational broadening and doppler shifting. Future work
    would need to develop ways to tune these extrinsic parameters alongside
    the fundamental stellar parameters.
    \item \textit{Framework Overhead}: As this algorithm is currently more 
    proof of concept than practical, it uses convenience functions from 
    various libraries, which naturally introduces some level of overhead and 
    leaves performance on the table. Future work would involve writing 
    custom functions expressly designed for \texttt{blase}, most likely a 
    complete rewrite of the library from the ground up.
    \item \textit{Pseudo-Interpretability}: Our algorithm boasts interpretability
    by considering spectral lines as the objects of interest as opposed to 
    the rather uninterpretable flux values of other approaches. However, this
    is only a step in the direction of interpretability. True interpretability
    would decompose a spectrum not into a set of spectral lines, but into a
    set of species component spectra, which requires a much more advanced 
    understanding of different species and their behavior, as well as
    direct access to a radiative transfer code as opposed to an off-the-shelf
    model grid. This approach would also extend the inference from just
    fundamental stellar parameters defined by a grid to any set of parameters
    accounted for in the radiative transfer model, down to specific species
    abundances.
    \item \textit{The Continuum Black Box}: Continuum normalization is a
    process that is not yet completely understood, and is currently done as 
    a preprocessing step with a fairly simple algorithm. Future work would
    dive deeper into the science of continuums and develop more advanced 
    methods that can discern continuums with greater accuracy and less
    modeling restrictions.
    \item \textit{One Voigt Fits All}: The current assumption of \texttt{blase}
    is that every spectral line is a Voigt profile. This assumption is largely
    true, but there are situations where that is simply not enough. Future
    studies need to account for more advanced spectral line profiles and 
    procedures to deal with phenomena such as ro-vibrational bands.
\end{enumerate}
In short, it is quite clear that \texttt{blase} is a long way from being a
powerhouse in spectral inference, however it represents a step down a road
not yet traveled, and the potential for future growth is immense. At the 
end of the day, we aim to develop a tool that can analyze spectrum based on 
how nature behaves rather than how data behaves.

\section*{Acknowledgements}
Text

\end{document}